---
title: Machine Learning Practical Tips
author: José Roberto Ayala Solares
date: '2018-04-30'
slug: ML-practical-tips
categories: []
tags:
  - notes
  - machine learning
summary: Practical tips from......
draft: true
---

[yomero](`r blogdown::shortcode("ref", "#probando")`) {#probando} 

## Feature Preprocessing

### Numeric features
Scaling: MinMax or Standardization (to make features' impacts on non-tree-based models roughly similar. Tree-based models do not depend on scaling).

Clipping (winsorization): for outliers

Rank transformation: set spaces between proper assorted values to be equal. It can be a better option than MinMax if we have outliers because rank transformation will move the outliers closer to other objects.

Log and Fractional power (i.e. sqrt) transformations: especially useful for neural networks. These can be useful because they drive 2 big values closer to the feature's average value. Also, values around zero become more distinguishable.

Sometimes it is useful to train a model on concatenated data frames produced by different preprocessings, or to mix models different preprocessed data

### Categorical and ordinal features
Label encoding: map each unique value to different numbers. Non-tree-based models require extra step by creating one-hot encoding.

Frequency encoding: maps categories to their frequencies. This preserves information about values' distribution and can help both linear and tree models. If two features have the same frequency, they become indistinguishable.

One-hot encoding: mainly used for non-tree-based models. It can introduce a lot of sparsity. To save memory, usage of sparse matrices can be benefitial if number of non-zero values is far less than half of total values.

## Feature Generation
It is powered by prior knowledge and EDA.

Feature interaction between different categorical features. 

## Datetime
1. Periodicity: second, minute, hour, day number in week, month, season, year
2. Time since either row-independent (a common date) or row-dependent (number of days left until next holiday / number of days passed since last holiday) event
3. Difference between dates

## Coordinates
1. Extract interesting places from the train/test data or additional data
2. Calculate distances to centers of clusters
3. Add aggregated statistics for surrounding data

# Missing data
## Missing value imputation
1. Replace NA with a value outside fixed value range (performance of linear models and NNs can suffer)
2. Add a binary feature that indicates missingness 
2. Replace with mean or median (beneficial for linear models and NN, but not for tree-based models)
3. Reconstruct the value

Be careful with imputed variables if you want to generate new features from them. Avoid replacing mising values before feature generation.

Outliers can be treated as missing values.

XGBoost and CatBoost can handle NA.


# Feature extraction from text and images


## For text
1. Bag of words
2. Word embeddings (word2vec)

Pipeline for bag of words (BOW):
* Preprocessing:
1. Lowercase
2. Lemmatization
3. Stemming
4. Stopwords

* BOW (Ngrams can help to use local context)

* Postprocessing:
1. Term frequency and Inverse document frequency (TFiDF) transformations


## For images
* Use pretrained models
* Data augmentation

It is crucial to understand the data generation process to set up a proper validation scheme.

Make distance plots (correlations, time that a feature is bigger than another feature).

Check that the data is shuffle: use mean and rolling mean of a feature or outcome variable.

```{r}
library(reticulate)
```

```{python, engine.path="/Applications/anaconda/bin/python"}
## sys.executable contains full path of the currently running Python interpreter. 
# import sys
# print(sys.executable)

import pandas as pd

titanic = pd.read_excel('http://biostat.mc.vanderbilt.edu/wiki/pub/Main/DataSets/titanic3.xls')

print titanic.head()
```

```{r}
#library(ggplot2)
#ggplot(py$titanic, aes(x = age)) +
#    geom_hist()
```

